Light inextensible strings under tension in the Schwarzschild geometry

### 2018-10-01 string_microarticle.tex
more copyediting

\received{1 May 2013}
\finalform{10 May 2013}
\accepted{13 May 2013}
\availableonline{15 May 2013}
\communicated{R. Hankin}

\begin{abstract}
%%%
Here I analyze the behaviour of light inextensible strings under
tension in the vicinity of a nonrotating black hole.  Proper length
(given by the spatial component of the Schwarzshild metric) is
minimized by the Euler-Lagrange equations.  Light inextensible taut
strings may assume topologically distinct configurations, and here I
give some numerical solutions and theoretical observations.
%%%%
\end{abstract}

@@

%% main text
\section{Introduction}
\label{intro}
In this short paper I consider the behaviour of a light inextensible
string in the exterior Schwarzschild metric.  A light string is not
subject to gravity, and being under tension means that its
configuration is an extremum of proper length.  Of course, no real
string is perfectly weightless or inextensible.  With tension~$T$,
gravitational acceleration $g$ and (proper) mass per (proper) unit
length~$\mu$, then ideal behaviour will be appropriate for
lengthscales $L\lesssim\frac{T}{\mu g}$.  Using Newtonian expressions
for the strength of gravity at the event horizon gives
$L\ll\frac{TGM}{\mu c^2}$.  For the string to behave as though light
on lengthscales comparable with the Schwarzschild radius would require
$T\gtrsim\mu c^2$.  Exceeding this criterion would violate the
dominant energy condition (and in this context we note that transverse
waves would propagate at speed~$\sqrt{T/\mu}\sim c$).  Nevertheless,
the idealization of a light inextensible string is an interesting and
informative limit: its analysis interprets the concept ``straight
line'' and suggests a possible way of manipulating a black hole.  The
Schwarzschild metric

@@

\begin{equation}
%  \int_{p_1}^{p_2}\sqrt{\frac{\left(r'\right)^2}{1-2M/r} + r^2}d\phi=
  \int_{p_1}^{p_2}\left(\frac{\left(r'\right)^2}{1-1/r} + r^2\right)^\frac{1}{2}d\phi=
  \int_{p_1}^{p_2}F\left(r,r'\right)d\phi.
\end{equation}

@@

\section{Results}
Light inextensible strings fall into two distinct classes: those with
two free ends, and those with one free end (the other ``end'' is
tangential to the event horizon).

Figure~\ref{closest_approach_non_self_intersecting} shows non-self
intersecting strings indexed by distance of approach to the
singularity, which is taken to occur at~$\phi=0$.  Strings with a
sufficiently close passage to the event horizon self-intersect;
numerical methods show that the critical value is~$\simeq
1.076109317$, with an uncertainty of 1 in the last digit: if the
closest approach is less than this value, the string crosses itself
at~$\phi=\pi/2$.  Figure~\ref{closest_approach_non_self_intersecting}
shows some non-self-intersecting strings and
figure~\ref{closest_approach_self_intersecting} shows some strings
that self-intersect.  Figure~\ref{one_free_end} shows some strings
that pass through~$\left(2,0\right)$ and are tangent to the event horizon.

\begin{figure}[h!]
\centering
\includegraphics[width=70mm]{closest_approach_nonselfintersecting.pdf}
\caption{Non-self intersecting strings arranged by distance~$d$ of
  closest approach to the event horizon, occurring at~$\phi=0$}
\label{closest_approach_non_self_intersecting}
\end{figure}

@@

The analysis above was restricted to the case~$r\geq 1$, where it is
possible to hold objects steady.  Of course, once within the event
horizon, even a light inextensible string cannot hold itself steady
(at least in Schwarzschild coordinates) and one would have to account
for the dynamics of the system.

In figure~\ref{closest_approach_non_self_intersecting}, one's
classical intuition would suggest that the black hole feels a force to
the right, and should start accelerating.  However, it is not clear by
what mechanism such a force would be transmitted: the situation is
steady, so no gravitational waves are produced; the string is
inextensible, so stores no energy.  Likewise,
Figure~\ref{one_free_end} would suggest that angular momentum is
transmitted to the black hole, but no mechanism is apparent.

>>>>

\received{day month year}
\finalform{day month year}
\accepted{day month year}
\availableonline{day month year}
\communicated{R. Hankin}

\begin{abstract}
%%%
Here I analyze the behaviour of stationary, light inextensible strings
under tension in the vicinity of a nonrotating black hole.  Proper
length (given by the spatial component of the Schwarzshild metric) is
minimized by the Euler-Lagrange equations.  Light inextensible taut
strings may assume topologically distinct configurations, and here I
give some numerical solutions and theoretical observations.  
%%%%
\end{abstract}

@@

%% main text
\section{Introduction}
\label{intro}
Light inextensible strings are a stalwart feature of elementary
physics.  A light string is not subject to gravity, and being under
tension means that its configuration is an extremum of proper length.
Of course, no real string is perfectly weightless or inextensible.
With tension~$T$, gravitational acceleration $g$ and (proper) mass per
(proper) unit length~$\mu$, then a string will behave as though light
on lengthscales $L\lesssim\frac{T}{\mu g}$.  Using Newtonian
expressions for the strength of gravity at the event horizon of a
black hole gives $L\lesssim\frac{TGM}{\mu c^2}$.  For the string to
behave as though light on lengthscales comparable with the
Schwarzschild radius would require $T\gtrsim\mu c^2$.  Exceeding this
criterion would violate the dominant energy condition (and in this
context we note that transverse waves would propagate at
speed~$\sqrt{T/\mu}\sim c$).  Nevertheless, the idealization of a
light inextensible string is an interesting and informative limit: its
analysis interprets the concept ``straight line'' and suggests a
possible way of manipulating a black hole.  The Schwarzschild metric

@@

\begin{equation}
%  \int_{p_1}^{p_2}\sqrt{\frac{\left(r'\right)^2}{1-2M/r} + r^2}d\phi=
  \int_{p_1}^{p_2}\left(\frac{\left(r'\right)^2}{1-1/r} + r^2\right)^\frac{1}{2}d\phi=
  \int_{p_1}^{p_2}F\left(r,r'\right)d\phi
\end{equation}

@@

\section{Results}
Solutions to equation~\ref{rdashdash} fall into two distinct classes:
those with two free ends, and those with one free end (the other
``end'' is tangential to the event horizon, in general).

Figure~\ref{closest_approach_non_self_intersecting} shows non-self
intersecting strings indexed by minimum distance of approach to the
singularity, which is taken to occur at~$\phi=0$.  Strings with a
sufficiently close passage to the event horizon self-intersect
(Figure~\ref{closest_approach_self_intersecting}).  Numerical methods
show that the critical value is~$\simeq 1.076109317$, with an
uncertainty of~1 in the last digit: if the closest approach is less
than this value, the string crosses itself at~$\phi=\pi/2$.
Figure~\ref{one_free_end} shows some strings that are tangent to the
event horizon and pass through the point~$\left(2,0\right)$

\begin{figure}[h!]
\centering
\includegraphics[width=70mm]{closest_approach_nonselfintersecting.pdf}
\caption{Non-self intersecting strings arranged by distance~$d$ of
  closest approach to the event horizon, occurring tangentially
  at~$\phi=0$}
\label{closest_approach_non_self_intersecting}
\end{figure}

@@

Light inextensible strings under tension are one interpretation of
``straight line'' in general relativity, the others being null
geodesics and world lines of massive objects.  Here it is shown that
such strings adopt non-trivial configurations including
self-intersecting paths.

In Figure~\ref{closest_approach_non_self_intersecting}, one's
classical intuition would suggest that the black hole feels a force to
the left, and should start accelerating.  However, it is not clear by
what mechanism such a force would be transmitted: the situation is
steady, so no gravitational waves are produced; the string is
inextensible, so stores no energy.  Likewise,
Figure~\ref{one_free_end} would suggest that angular momentum is
transmitted to the black hole, but no mechanism is apparent.
